---
title: "Project 3: STAT302package Tutorial"
author: "Judy Tian"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This is a R package including four functions: T-test function `my_t.test`, Linear Model function `my_lm`, Random Forest Cross-Validation function `my_knn_cv` and K-Nearest Neighbors Cross-Validation function `my_rf_cv`.

You can install the package through GitHub using:

``` {r, eval = FALSE}
devtools::install_github("JudyTi/STAT302package")
```

To begin, please load the package:

```{r setup}
library(STAT302package)
data("my_gapminder")
```

To view the vignette:

``` {r, eval = FALSE}
devtools::install_github("JudyTi/STAT302package", build_vignette = TRUE, build_opts = c())
library(STAT302package)
# Use this to view the vignette in the STAT302package HTML help
help(package = "STAT302package", help_type = "html")
# Use this to view the vignette as an isolated HTML file
utils::browseVignettes(package = "STAT302package")
```

## Tutorial for `my_t.test`

This function performs a one sample t-test in R. 

To demonstrate and present the function, the following three tests will use the `lifeExp` data from `my_gapminder` with the p-value cut-off of $\alpha = 0.05$.

### Two side test

The first hypothesis test is:
\[\begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}\]
  
This is saying that the default null hypothesis $H_0$ is `mu` which equals to 60 and the alternative hypothesis $H_a$ is `mu` which isn't equal to 60. It's an example of two side test.

For this two side test, we could use the following code:

``` {r}
# two-sided t-test
two_side <- my_t.test(my_gapminder$lifeExp, alternative = "two.sided", mu = 60)
```
  
Thus the p value of this test is `r two_side$p_val`. This is greater than our pre-determined cut-off of 0.05, so we conclude that the result isn't statistically significant. Therefore, we don't reject null hypothesis $H_0$.

### One side test -- Upper tail

The second hypothesis test is:
\[\begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}\]

This is saying that the default null hypothesis $H_0$ is `mu` which equals to 60 and the alternative hypothesis $H_a$ is `mu` which not equals to 60. It's an example of two side test.

For this one side upper tail test, we could use the following code:

``` {r}
# one-sided t-test -- upper tail
upper_test <- my_t.test(my_gapminder$lifeExp, alternative = "greater", mu = 60)
```

Thus the p value of this test is `r upper_test$p_val`. This is greater than our pre-determined cut-off of 0.05, so we conclude that the result isn't statistically significant. Therefore, we don't reject null hypothesis $H_0$.

### One side test -- Lower tail

The third hypothesis test is:
\[\begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}\]

This is saying that the default null hypothesis $H_0$ is `mu` which equals to 60 and the alternative hypothesis $H_a$ is `mu` which isn't equal to 60. It's an example of two side test.

For this one side lower tail test, we could use the following code:

``` {r}
# one-sided t-test -- lower tail
lower_test <- my_t.test(my_gapminder$lifeExp, alternative = "less", mu = 60)
```

Thus the p value of this test is `r lower_test$p_val`. This is less than our pre-determined cut-off of 0.05, so we conclude that the result is statistically significant. Therefore, we reject null hypothesis $H_0$.


## Tutorial for `my_lm`

## Tutorial for `my_knn_cv` using `my_penguins`

## Tutorial for `my_rf_cv`
